Sakura Blossom Prediction Model for Japan

### 2024-11-28 paper/paper.qmd
need fix mirror issue of words

# Display the sample as a clean table
kable(sakura_sample, 
      col.names = c("ID", "Location", "Latitude", "Longitude", "Year", "Month",
                    "Flower Day", "Full Bloom Day", "Mean Temp"))
```
For historical sakura data, since the earliest data in this dataset can be traced back to 812, however the accuracy of temperature measurements in the early years is questionable, and data being recorded only in the Kyoto area, there may be some unavoidable bias. Therefore, when building the prediction model, we will only use modern sakura data for fitting, and historical data will only serve as a comparison to help us understand the historical situation.

@@

@@

### Fixed Effects Coefficients with 95% Credible Intervals

The @fig-fix highlights the relationships between key predictors and sakura flowering dates based on the Bayesian hierarchical model. The intercept -47.4751 represents the expected flowering date when all predictors are zero, though it’s mainly a baseline without much practical interpretation since predictors like temperature and location are scaled.

@@

Longitude, in contrast, has a much smaller and less certain effect. The coefficient -0.2695 suggests that flowering may occur earlier as longitude increases, but the credible interval [-0.7217, 0.1194] includes zero. This uncertainty means longitude’s impact is less clear, and it likely plays a minor role compared to temperature and latitude.

```{r}
#| label: fig-fix   # For citation in Quarto
#| tbl-cap: "Fixed Effects Coefficients with 95% Credible Intervals"  # Caption for the table
#| echo: false  # Suppress code in output
#| warning: false 
#| message: false
#| fig-align: center

@@

### Performance Metrics for Random Effects (Year and Region)

@@

@@

In @fig-ppcheckandposteriorvsprior we implement a posterior predictive check. 
This shows the density overlay plot demonstrates a strong alignment between the observed sakura flowering dates and the model's posterior predictions. The observed density, represented by the dark blue line, closely matches the predictive densities (lighter blue lines) across most of the distribution, indicating that the model captures the central tendency and major patterns effectively. The primary peak around 80 is well-represented, showing the model's strength in predicting the most common flowering dates. Additionally, the secondary peak around 120 aligns reasonably well, suggesting the model's ability to handle more complex patterns, such as bimodal distributions. However, slight discrepancies are observed in the tails of the distribution, particularly below 40 and above 140, where the predictive densities deviate from the observed data. These divergences indicate potential challenges in capturing extreme values or outliers, which may be addressed by including additional predictors, such as microclimatic conditions or species-specific traits.

>>>>

# Display the sample as a clean table
kable(sakura_sample, 
      col.names = c("ID", "Location", "Latitude", "Longitude", "Year", "Month",
                    "Flower Day", "Full Bloom Day", "Mean Temp"))
```
For historical sakura data is provided by Prof. Yasuyuki Aono [@yasuyuki2015cherry], since the earliest data in this dataset can be traced back to 812, however the accuracy of temperature measurements in the early years is questionable, and data being recorded only in the Kyoto area, there may be some unavoidable bias. Therefore, when building the prediction model, we will only use modern sakura data for fitting, and historical data will only serve as a comparison to help us understand the historical situation.

@@

### Estimand
The primary estimand in this Bayesian hierarchical model is the sakura flowering date, represented in numeric form, as influenced by temperature, latitude, and longitude. The model estimates both the fixed effects of these predictors and the random effects associated with yearly and regional variations, capturing localized deviations and inter-annual trends. Additionally, the model estimates residual variance to account for unexplained variability, ensuring a comprehensive understanding of the factors driving sakura phenology. These estimands enable the quantification of the relationship between climatic and geographic predictors and the timing of sakura blooming, while accounting for spatial and temporal heterogeneity.

@@

### Posterior: Fixed Effects Coefficients with 95% Credible Intervals

The @tbl-fix highlights the relationships between key predictors and sakura flowering dates based on the Bayesian hierarchical model and the distribution of posterior are shown in @fig-pos-dis. The intercept -47.4751 represents the expected flowering date when all predictors are zero, though it’s mainly a baseline without much practical interpretation since predictors like temperature and location are scaled.

@@

Longitude, in contrast, has a much smaller and less certain effect. The coefficient -0.2695 suggests that flowering may occur earlier as longitude increases, but the credible interval [-0.7217, 0.1194] includes zero. This uncertainty means longitude’s impact is less clear, and it likely plays a minor role compared to temperature and latitude.
```{r}
#| label: fig-pos-dis
#| fig-cap: "Posterior Distributions"
#| echo: false
#| warning: false 
#| message: false
#| fig-align: center

# Extract posterior samples for fixed effects
posterior_samples <- as.data.frame(flower_model)

# Select relevant predictors from the posterior samples
predictor_columns <- c("(Intercept)", "latitude", "longitude", "month_mean_temp")
posterior_subset <- posterior_samples[, predictor_columns]

# Transform data into long format for ggplot
posterior_long <- posterior_subset %>%
  pivot_longer(
    cols = everything(),
    names_to = "Predictor",
    values_to = "Estimate"
  )

# Create the posterior distribution plot
ggplot(posterior_long, aes(x = Estimate, fill = Predictor)) +
  geom_density(alpha = 0.7) +
  facet_wrap(~ Predictor, scales = "free", ncol = 2) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "red") +
  labs(
    title = "Posterior Distributions of Coefficients",
    x = "Coefficient Estimate",
    y = "Density"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5), # Center the title
    legend.position = "none"
  )

```{r}
#| label: tbl-fix   # For citation in Quarto
#| tbl-cap: "Fixed Effects Coefficients with 95% Credible Intervals"  # Caption for the table
#| echo: false  # Suppress code in output
#| warning: false 
#| message: false
#| fig-align: center

@@

### Posterior: Performance Metrics for Random Effects (Year and Region)

@@

## Survey design and sampling techniques

@@

In @fig-ppcheckandposteriorvsprior we implement a posterior predictive check and the distribution of posterior are show in @fig-pos-dis. 
This shows the density overlay plot demonstrates a strong alignment between the observed sakura flowering dates and the model's posterior predictions. The observed density, represented by the dark blue line, closely matches the predictive densities (lighter blue lines) across most of the distribution, indicating that the model captures the central tendency and major patterns effectively. The primary peak around 80 is well-represented, showing the model's strength in predicting the most common flowering dates. Additionally, the secondary peak around 120 aligns reasonably well, suggesting the model's ability to handle more complex patterns, such as bimodal distributions. However, slight discrepancies are observed in the tails of the distribution, particularly below 40 and above 140, where the predictive densities deviate from the observed data. These divergences indicate potential challenges in capturing extreme values or outliers, which may be addressed by including additional predictors, such as microclimatic conditions or species-specific traits.
